---
title: "LAI analysis of the core Vietnamese Robusta collection"
author: "Tram"
date: "6/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(LEA)
library(Rsamtools)
library(tidyverse)
library(viridis)
library(Rcpp)
library(data.table)
library(karyoploteR)
library(gt)
library(gridExtra)
library(ggpubr)
library(S4Vectors)
library(aplot)
```

# copy elai results from IFB to Itrop
```{bash}
rsync -vaurP baotram@core.cluster.france-bioinformatique.fr:/shared/projects/elai_most/data/snakelai_out_all/elai_results /data/projects/rob/data/most_elai/elai_out/
rsync -vaurP baotram@core.cluster.france-bioinformatique.fr:/shared/projects/elai_most/data/snakelai_out_all/source /data/projects/rob/data/most_elai/elai_out/
rsync -vaurRP baotram@core.cluster.france-bioinformatique.fr:/shared/projects/elai_most/data/snakelai_out_all/./pos/CC1.8.Chr01/even/test_1.pos /data/projects/rob/data/most_elai/elai_out/
```

```{r}
elai_outdir <- "/data/projects/rob/data/most_elai/elai_out"
source_dir <- file.path(elai_outdir, "source")
plot_dir <- "/data/projects/rob/data/most_elai/plots"
plot_dir <- "/shared/projects/elai_most/diversity_wasi/plot_paper"
```

# list of individuals
```{r}
# vcf_file <- list.files(source_dir, "CC1.8.Chr07.recode.vcf", recursive = T, full.names = T)
vcf_header <- scanBcfHeader("/data/projects/rob/data/most_snp_2021/all_final_thin5000_maf002.recode.vcf")
vcf_names <- vcf_header[[1]]@listData$Sample
```

```{r}
info_file <- "/data/projects/rob/data/most_elai/snps_data/sequence_info_final.tsv"
ind_info <- read.delim(info_file)
ind_names <- sapply(ind_names, function(i) 
  ifelse(any(grepl(i, ind_info$Label)), 
         grep(i, ind_info$Label, value = T),
         i),
  USE.NAMES = F)
```



# combine elai sets


```{r}
all_chr <- GRanges()
all_chr <- do.call(GRangesList, sapply(1:11, function(i) {
  chr <- readRDS(paste0("/data/projects/rob/data/most_elai/merged_elai/chrom_", sprintf("%02d", i), ".RDS"))
  return(chr)
}, simplify = T)) %>% unlist
seqlevels(all_chr) <- gsub("CC1.8.Chr", "Chr ", seqlevels(all_chr))
all_chr$individual <- gsub(".", "-", all_chr$individual, fixed = T)
all_chr$individual <- gsub("m", "", all_chr$individual, fixed = T)
```


```{r}
all_windows_large <- all_chr[width(all_chr) > 1e6,]
```

```{r}
group_col <- c("#30123BFF", "#28BBECFF", "#A2FC3CFF", "#FB8022FF", "#7A0403FF")
names(group_col) <- c("ER", "OB", "C", "AG", "D")

# genome
genome_fa <- "/data3/projects/vietcaf/reference/CC1.8_v2_pseudomolecule_cat.fa"
genome <- scanFa(genome_fa)
genome <- genome[grepl("Chr", names(genome))] # remove contigs
names(genome) <- gsub(" .+", "", names(genome))
names(genome) <- gsub("CC1.8.Chr", "Chr ", names(genome))
genome_GR <- GRanges(seqnames = names(genome), ranges = IRanges(start = 1, end = width(genome)))
```


```{r}
plot_dir <- "./plots"
# plot each individual separately
for (id in unique(all_chr$individual)) {
  all_windows_ind <- all_chr[all_chr$individual == id]
  all_windows_large <- all_windows_ind[width(all_windows_ind) > 1e6,]
  # all_windows_small <- all_windows_ind[width(all_windows_ind) < 1e6,]
  groups <- colnames(mcols(all_chr))[colnames(mcols(all_chr)) != "individual"]
  
  pp1 <- getDefaultPlotParams(plot.type=6)
  pp1$data1height <- 5
  pp1$data2height <- 5
  pp1$topmargin <- 20
  pp1$bottommargin <- 20
  pp1$leftmargin <- .07
  pp1$rightmargin <- .15
  
  tiff(file.path(plot_dir, paste0(gsub("\\.", "_", id), ".tiff")),
     width = 11, heigh = 6, unit = "in", res = 1000)
  par(mar = c(2,2,2,2))
  gkp <- plotKaryotype(genome = genome_GR, plot.type = 6,
                       plot.params = pp1, main = gsub("\\.", "-", id)) 
  kpDataBackground(gkp, data.panel ="ideogram", color = "grey90")
  kpAddBaseNumbers(gkp, tick.dist = 1e7)
  for (g in 1:length(groups)) {
    if (g == 1) {
      ga <- 0
      gz <- mcols(all_windows_large)[,groups[g]]
    } else {
      ga <- gz
      gz <- gz+mcols(all_windows_large)[,groups[g]]
    }
    kpBars(gkp, data = all_windows_large, y0 = ga, y1 = gz, col = group_col[groups[g]], border = NA)
  }
  kpAxis(gkp, side = 2, data.panel=1, cex = 0.7)
  legend("right", #x = 0.9, y = 0.4, #"bottomright", #  yjust = .5, xjust = 0, # 
         title = "ancestry", legend = c(gsub("group_", "", names(group_col)), "undetermined"), 
         fill = c(group_col, "grey90"), bty = "n", xjust = 0)
  dev.off()
}



```


## summary table for overall proportion
```{r}
## summary table for overall proportion

global <- do.call(rbind, lapply(unique(all_windows_large$individual), function(i) {
  # cat("> individual:", i, "\n")
  all_windows_ind <- all_windows_large[all_windows_large$individual == i,]
  all_windows_ind <- all_windows_ind[width(all_windows_ind) > 1e6]
  grps <- mcols(all_windows_ind) %>% as.data.frame() %>% dplyr::select(-individual) %>% colnames()
  glob_adm <- do.call(rbind, lapply(grps, function(g) {
    pw <- sum(mcols(all_windows_ind)[,g]*width(all_windows_ind))/sum(width(genome_GR))
    return(pw)
  }))
  glob_adm <- as.data.frame(t(glob_adm))
  colnames(glob_adm) <- grps
  rownames(glob_adm) <- i
  return(glob_adm)
}))

global <- global[stringr::str_sort(rownames(global), numeric = T), stringr::str_sort(colnames(global))]
undetermined <- 1 - rowSums(global)
global$undetermined <- undetermined
colnames(global) <- gsub("group_", "ancestry ", colnames(global))
rownames(global) <- gsub("m", "", rownames(global))
rownames(global) <- gsub("\\.", "-", rownames(global))

t <- global %>% 
  as.data.frame() %>% 
  gt(rownames_to_stub = T) %>% 
  fmt_number(columns = everything(), decimals = 3) %>% 
  data_color(colors = scales::col_numeric(palette = c("white", group_col["AG"]), domain = c(0,1)),
             columns = "AG") %>% 
  data_color(colors = scales::col_numeric(palette = c("white", group_col["ER"]), domain = c(0,1)),
             columns = "ER") %>% 
  data_color(colors = scales::col_numeric(palette = c("white", group_col["OB"]), domain = c(0,1)),
             columns = "OB") %>% 
  data_color(colors = scales::col_numeric(palette = c("white", group_col["D"]), domain = c(0,1)),
             columns = "D") %>% 
  data_color(colors = scales::col_numeric(palette = c("white", group_col["C"]), domain = c(0,1)),
             columns = "C") %>% 
  data_color(colors = scales::col_numeric(palette = c("white", "black"), domain = c(0,1)),
             columns = "undetermined") 
gtsave(t, file.path(plot_dir, "global_ancestry_elai.pdf"))
gt <- as_ggplot(t, vwidth = 2400, vheight = 2000)
```

```{r}
write.table(global, "global_lai.tsv", sep = "\t")
```

## correlation between lai and snmf global inference
```{r}
snmfPrj <- "./snmf/thin_all.snmfProject"
all_snmf <- load.snmfProject(snmfPrj)
k <- 5
snmf_prop <- Q(all_snmf, k, 
               run = which.min(cross.entropy(all_snmf, k)))
colnames(snmf_prop) <- c("C", "OB", "ER", "AG", "D")
snmf_glob <- snmf_prop %>% 
  as.data.frame() %>% 
  mutate(id = gsub("m$", "", vcf_names)) %>% 
  filter(grepl("(S-|TR)", id)) %>% 
  pivot_longer(cols = -id, names_to = "ancestry", values_to = "snmf")

elai_glob <- global %>% 
  rownames_to_column("id") %>% 
  select(-undetermined) %>% 
  pivot_longer(cols = -id, names_to = "ancestry", values_to = "elai")

merge(snmf_glob, elai_glob) %>% 
  group_by(id) %>%
  summarise(id_correlation = cor(snmf, elai, method = "pearson")) %>% 
  summarise(correlation = mean(id_correlation)) ## result: 0.972
```


## plot all genomes admixed with AG > 10%

```{r}
# plot all genomes admixed with AG > 10%
pp2 <- getDefaultPlotParams(plot.type=7)
pp2$data1outmargin <- 0
# pp2$data1height <- 5
# pp2$data2height <- 5
pp2$topmargin <- 40
# pp2$bottommargin <- 20
pp2$leftmargin <- .07
pp2$rightmargin <- .2
pp2$ideogramlateralmargin <- 0.003

inds_AG <- rownames(global)[global$AG > .1 & global$D < 0.1 & global$OB < 0.03]
# inds_AG <- gsub("-", "\\.", inds_AG)
# inds_AG <- gsub("TR6", "TR6m", inds_AG)


tiff(file.path(plot_dir, "all_AG.tiff"),
     width = 20, heigh = 6, unit = "in", res = 1000)
par(mar = c(2,2,2,2))
gkp <- plotKaryotype(genome = genome_GR, plot.type = 7,
                     plot.params = pp2)
kpDataBackground(gkp, data.panel ="ideogram", color = "white")
kpAddBaseNumbers(gkp, tick.dist = 1e7)
for (id in 1:length(inds_AG)) {
  if (id == 1) {
    ra <- 0
    rz <- 1/length(inds_AG)-.02
  } else {
    ra <- rz +.02
    rz <- rz + 1/length(inds_AG)
  }
  ind_lab <- gsub("m", "", inds_AG[id])
  ind_lab <- gsub("\\.", "-", ind_lab)
  chr_ind <- all_chr[all_chr$individual == inds_AG[id]]
  all_windows_large <- chr_ind[width(chr_ind) > 1e6,]
  all_windows_small <- chr_ind[width(chr_ind) < 1e6,]
  groups <- colnames(mcols(all_chr))[colnames(mcols(all_chr)) != "individual"]
  
  kpBars(gkp, data = genome_GR, y0 = rep(0, length(genome_GR)), y1 = rep(1, length(genome_GR)), col = "gray80", border = NA, r0 = ra, r1 = rz)
  kpAxis(gkp, side = 1, data.panel=1, cex = 0.7, r0 = ra, r1 = rz)
  kpText(gkp, chr=seqnames(genome_GR)[length(genome_GR)], x=width(genome_GR)[length(genome_GR)] + 2.6e7, y=0.5, labels = ind_lab, r0 = ra, r1 = rz)
  for (g in 1:length(groups)) {
    if (g == 1) {
      ga <- 0
      gz <- mcols(all_windows_large)[,groups[g]]
    } else {
      ga <- gz
      gz <- gz+mcols(all_windows_large)[,groups[g]]
    }
    kpBars(gkp, data = all_windows_large, y0 = ga, y1 = gz, col = group_col[groups[g]], border = NA, r0 = ra, r1 = rz)
  }
  # kpPlotRegions(gkp, data = all_windows_large, data.panel = "ideogram", col = "grey20", avoid.overlapping = T, r0 = ra, r1 = rz)
  # kpPlotRegions(gkp, data = all_windows_small, data.panel = "ideogram", col = "orange3", avoid.overlapping = T)
}
legend("right", #x = 0.9, y = 0.4, #"bottomright", #  yjust = .5, xjust = 0, # 
       title = "ancestry", legend = c(gsub("group_", "", names(group_col)), "undetermined"), 
       fill = c(group_col, "grey90"), bty = "n", xjust = 0, cex = 1.2)
dev.off()
```

```{r}
adm_blocks %>% 
  pivot_longer(cols = OB:ER, names_to = "ancestry", values_to = "proportion") %>% 
  mutate(start = start/1e6,
         end = end/1e6,
         individual = factor(individual, 
                             levels = adm_order)) %>% 
  group_by(chromosome, start, end, individual) %>% 
  mutate(proportion_min = cum_min(proportion),
         proportion_max = cum_max(proportion)) %>% 
  filter(individual %in% inds_AG) %>% 
  ggplot() + 
  facet_grid(rows = vars(individual), cols = vars(chromosome),
             scale = "free_x", space = "free_x", switch = "x", shrink = F) + 
  geom_rect(aes(xmin = start, xmax = end,
                ymin = proportion_min, ymax = proportion_max,
                fill = ancestry),
            show.legend = F) +
  scale_fill_manual(values = group_col, na.value = "gray") +
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        axis.ticks = element_line(),
        axis.text.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        # strip.background.y = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.placement = "outside",
        strip.text.y = element_text(angle = 0, h = 0),
        panel.spacing.x = unit(4, "pt"), 
        panel.spacing.y = unit(0.5, "pt"), 
        panel.background = element_rect(fill = "gray60", color = "white")) +
  scale_y_continuous(breaks = 0.5) +
  scale_x_continuous(breaks = seq(0, 80, 10), expand = c(0, 0)) +
  xlab("Position (Mb)") + ylab("Ancestry assignment")
```


## plot all genomes admixed with D > 10%

```{r}
# plot all genomes admixed with D > 10%

inds_D <- rownames(global)[global$D > .1 & global$AG == 0]
# inds_D <- gsub("-", "\\.", inds_D)


tiff(file.path(plot_dir, "all_D.tiff"),
     width = 20, heigh = 12, unit = "in", res = 1000)
par(mar = c(2,2,2,2))
gkp <- plotKaryotype(genome = genome_GR, plot.type = 7,
                     plot.params = pp2)
kpDataBackground(gkp, data.panel ="ideogram", color = "white")
kpAddBaseNumbers(gkp, tick.dist = 1e7)
for (id in 1:length(inds_D)) {
  if (id == 1) {
    ra <- 0
    rz <- 1/length(inds_D)-.02
  } else {
    ra <- rz +.02
    rz <- rz + 1/length(inds_D)
  }
  ind_lab <- gsub("m", "", inds_D[id])
  ind_lab <- gsub("\\.", "-", ind_lab)
  chr_ind <- all_chr[all_chr$individual == inds_D[id]]
  all_windows_large <- chr_ind[width(chr_ind) > 1e6,]
  all_windows_small <- chr_ind[width(chr_ind) < 1e6,]
  groups <- colnames(mcols(all_chr))[colnames(mcols(all_chr)) != "individual"]
  
  kpBars(gkp, data = genome_GR, y0 = rep(0, length(genome_GR)), y1 = rep(1, length(genome_GR)), col = "gray80", border = NA, r0 = ra, r1 = rz)
  kpAxis(gkp, side = 1, data.panel=1, cex = 0.7, r0 = ra, r1 = rz)
  kpText(gkp, chr=seqnames(genome_GR)[length(genome_GR)], x=width(genome_GR)[length(genome_GR)] + 2.6e7, y=0.5, labels = ind_lab, r0 = ra, r1 = rz)
  for (g in 1:length(groups)) {
    if (g == 1) {
      ga <- 0
      gz <- mcols(all_windows_large)[,groups[g]]
    } else {
      ga <- gz
      gz <- gz+mcols(all_windows_large)[,groups[g]]
    }
    kpBars(gkp, data = all_windows_large, y0 = ga, y1 = gz, col = group_col[groups[g]], border = NA, r0 = ra, r1 = rz)
  }
  # kpPlotRegions(gkp, data = all_windows_large, data.panel = "ideogram", col = "grey20", avoid.overlapping = T, r0 = ra, r1 = rz)
  # kpPlotRegions(gkp, data = all_windows_small, data.panel = "ideogram", col = "orange3", avoid.overlapping = T)
}
legend("right", #x = 0.9, y = 0.4, #"bottomright", #  yjust = .5, xjust = 0, # 
       title = "ancestry", legend = c(gsub("group_", "", names(group_col)), "undetermined"), 
       fill = c(group_col, "grey90"), bty = "n", xjust = 0, cex = 1.2)
dev.off()
```

```{r}
adm_blocks %>% 
  pivot_longer(cols = OB:ER, names_to = "ancestry", values_to = "proportion") %>% 
  mutate(start = start/1e6,
         end = end/1e6,
         individual = factor(individual, 
                             levels = adm_order)) %>% 
  group_by(chromosome, start, end, individual) %>% 
  mutate(proportion_min = cum_min(proportion),
         proportion_max = cum_max(proportion)) %>% 
  filter(individual %in% inds_D) %>% 
  ggplot() + 
  facet_grid(rows = vars(individual), cols = vars(chromosome),
             scale = "free_x", space = "free_x", switch = "x", shrink = F) + 
  geom_rect(aes(xmin = start, xmax = end,
                ymin = proportion_min, ymax = proportion_max,
                fill = ancestry),
            show.legend = F) +
  scale_fill_manual(values = group_col, na.value = "gray") +
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        axis.ticks = element_line(),
        axis.text.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        # strip.background.y = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.placement = "outside",
        strip.text.y = element_text(angle = 0, h = 0),
        panel.spacing.x = unit(4, "pt"), 
        panel.spacing.y = unit(0.5, "pt"), 
        panel.background = element_rect(fill = "gray60", color = "white")) +
  scale_y_continuous(breaks = 0.5) +
  scale_x_continuous(breaks = seq(0, 80, 10), expand = c(0, 0)) +
  xlab("Position (Mb)") + ylab("Ancestry assignment")
```

## plot all genomes pure E
```{r}
# plot all genomes pure E

inds_ER <- rownames(global)[global$AG == 0 & global$D == 0 & global$OB == 0]
# inds_ER <- gsub("-", "\\.", inds_ER)
# inds_ER <- ifelse(grepl("TR", inds_ER), paste0(inds_ER, "m"), inds_ER)


tiff(file.path(plot_dir, "all_ER.tiff"),
     width = 20, heigh = 8, unit = "in", res = 1000)
par(mar = c(2,2,2,2))
gkp <- plotKaryotype(genome = genome_GR, plot.type = 7,
                     plot.params = pp2)
kpDataBackground(gkp, data.panel ="ideogram", color = "white")
kpAddBaseNumbers(gkp, tick.dist = 1e7)
for (id in 1:length(inds_ER)) {
  if (id == 1) {
    ra <- 0
    rz <- 1/length(inds_ER)-.02
  } else {
    ra <- rz +.02
    rz <- rz + 1/length(inds_ER)
  }
  ind_lab <- gsub("m", "", inds_ER[id])
  ind_lab <- gsub("\\.", "-", ind_lab)
  chr_ind <- all_chr[all_chr$individual == inds_ER[id]]
  all_windows_large <- chr_ind[width(chr_ind) > 1e6,]
  all_windows_small <- chr_ind[width(chr_ind) < 1e6,]
  groups <- colnames(mcols(all_chr))[colnames(mcols(all_chr)) != "individual"]
  
  kpBars(gkp, data = genome_GR, y0 = rep(0, length(genome_GR)), y1 = rep(1, length(genome_GR)), col = "gray80", border = NA, r0 = ra, r1 = rz)
  kpAxis(gkp, side = 1, data.panel=1, cex = 0.7, r0 = ra, r1 = rz)
  kpText(gkp, chr=seqnames(genome_GR)[length(genome_GR)], x=width(genome_GR)[length(genome_GR)] + 2.6e7, y=0.5, labels = ind_lab, r0 = ra, r1 = rz)
  for (g in 1:length(groups)) {
    if (g == 1) {
      ga <- 0
      gz <- mcols(all_windows_large)[,groups[g]]
    } else {
      ga <- gz
      gz <- gz+mcols(all_windows_large)[,groups[g]]
    }
    kpBars(gkp, data = all_windows_large, y0 = ga, y1 = gz, col = group_col[groups[g]], border = NA, r0 = ra, r1 = rz)
  }
  # kpPlotRegions(gkp, data = all_windows_large, data.panel = "ideogram", col = "grey20", avoid.overlapping = T, r0 = ra, r1 = rz)
  # kpPlotRegions(gkp, data = all_windows_small, data.panel = "ideogram", col = "orange3", avoid.overlapping = T)
}
legend("right", #x = 0.9, y = 0.4, #"bottomright", #  yjust = .5, xjust = 0, # 
       title = "ancestry", legend = c(gsub("group_", "", names(group_col)), "undetermined"), 
       fill = c(group_col, "grey90"), bty = "n", xjust = 0, cex = 1.2)
dev.off()


```



## plot all the other genomes

```{r}
# plot all the other genomes

# inds_ER <- rownames(global)[global$AG == 0 & global$D == 0 & global$OB == 0]
# inds_D <- rownames(global)[global$D > .1]
# inds_AG <- rownames(global)[global$AG > .1]
inds_other <- rownames(global)[!rownames(global) %in% c (inds_ER, inds_D, inds_AG)]
# inds_other <- gsub("-", "\\.", inds_other)
# inds_other <- ifelse(grepl("TR", inds_other), paste0(inds_other, "m"), inds_other)


tiff(file.path(plot_dir, "all_others.tiff"),
     width = 20, heigh = 16, unit = "in", res = 1000)
par(mar = c(2,2,2,2))
gkp <- plotKaryotype(genome = genome_GR, plot.type = 7,
                     plot.params = pp2)
kpDataBackground(gkp, data.panel ="ideogram", color = "white")
kpAddBaseNumbers(gkp, tick.dist = 1e7)
for (id in 1:length(inds_other)) {
  if (id == 1) {
    ra <- 0
    rz <- 1/length(inds_other)-.02
  } else {
    ra <- rz +.02
    rz <- rz + 1/length(inds_other)
  }
  ind_lab <- gsub("m", "", inds_other[id])
  ind_lab <- gsub("\\.", "-", ind_lab)
  chr_ind <- all_chr[all_chr$individual == inds_other[id]]
  all_windows_large <- chr_ind[width(chr_ind) > 1e6,]
  all_windows_small <- chr_ind[width(chr_ind) < 1e6,]
  groups <- colnames(mcols(all_chr))[colnames(mcols(all_chr)) != "individual"]
  
  kpBars(gkp, data = genome_GR, y0 = rep(0, length(genome_GR)), y1 = rep(1, length(genome_GR)), col = "gray80", border = NA, r0 = ra, r1 = rz)
  kpAxis(gkp, side = 1, data.panel=1, cex = 0.7, r0 = ra, r1 = rz)
  kpText(gkp, chr=seqnames(genome_GR)[length(genome_GR)], x=width(genome_GR)[length(genome_GR)] + 2.6e7, y=0.5, labels = ind_lab, r0 = ra, r1 = rz)
  for (g in 1:length(groups)) {
    if (g == 1) {
      ga <- 0
      gz <- mcols(all_windows_large)[,groups[g]]
    } else {
      ga <- gz
      gz <- gz+mcols(all_windows_large)[,groups[g]]
    }
    kpBars(gkp, data = all_windows_large, y0 = ga, y1 = gz, col = group_col[groups[g]], border = NA, r0 = ra, r1 = rz)
  }
  # kpPlotRegions(gkp, data = all_windows_large, data.panel = "ideogram", col = "grey20", avoid.overlapping = T, r0 = ra, r1 = rz)
  # kpPlotRegions(gkp, data = all_windows_small, data.panel = "ideogram", col = "orange3", avoid.overlapping = T)
}
legend("right", #x = 0.9, y = 0.4, #"bottomright", #  yjust = .5, xjust = 0, # 
       title = "ancestry", legend = c(gsub("group_", "", names(group_col)), "undetermined"), 
       fill = c(group_col, "grey90"), bty = "n", xjust = 0, cex = 1.2)
dev.off()
```

```{r}
adm_blocks %>% 
  pivot_longer(cols = OB:ER, names_to = "ancestry", values_to = "proportion") %>% 
  mutate(start = start/1e6,
         end = end/1e6,
         individual = factor(individual, 
                             levels = adm_order)) %>% 
  group_by(chromosome, start, end, individual) %>% 
  mutate(proportion_min = cum_min(proportion),
         proportion_max = cum_max(proportion)) %>% 
  filter(individual %in% inds_other[!inds_other %in% ind_3adm]) %>% 
  ggplot() + 
  facet_grid(rows = vars(individual), cols = vars(chromosome),
             scale = "free_x", space = "free_x", switch = "x", shrink = F) + 
  geom_rect(aes(xmin = start, xmax = end,
                ymin = proportion_min, ymax = proportion_max,
                fill = ancestry),
            show.legend = F) +
  scale_fill_manual(values = group_col, na.value = "gray") +
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        axis.ticks = element_line(),
        axis.text.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        # strip.background.y = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.placement = "outside",
        strip.text.y = element_text(angle = 0, h = 0),
        panel.spacing.x = unit(4, "pt"), 
        panel.spacing.y = unit(0.5, "pt"), 
        panel.background = element_rect(fill = "gray60", color = "white")) +
  scale_y_continuous(breaks = 0.5) +
  scale_x_continuous(breaks = seq(0, 80, 10), expand = c(0, 0)) +
  xlab("Position (Mb)") + ylab("Ancestry assignment")
```

## plot only 3-way admixed individuals
```{r}
ind_3adm <- c("S-70", "S-134", "S-105", "S-122")

tiff(file.path(plot_dir, "ind_3way_adm.tiff"),
     width = 15, heigh = 6, unit = "in", res = 1000)
par(mar = c(2,2,2,2))
gkp <- plotKaryotype(genome = genome_GR, plot.type = 7,
                     plot.params = pp2)
kpDataBackground(gkp, data.panel ="ideogram", color = "white")
kpAddBaseNumbers(gkp, tick.dist = 1e7)
for (id in 1:length(ind_3adm)) {
  if (id == 1) {
    ra <- 0
    rz <- 1/length(ind_3adm)-.02
  } else {
    ra <- rz +.02
    rz <- rz + 1/length(ind_3adm)
  }
  ind_lab <- gsub("m", "", ind_3adm[id])
  ind_lab <- gsub("\\.", "-", ind_lab)
  chr_ind <- all_chr[all_chr$individual == ind_3adm[id]]
  all_windows_large <- chr_ind[width(chr_ind) > 1e6,]
  all_windows_small <- chr_ind[width(chr_ind) < 1e6,]
  groups <- colnames(mcols(all_chr))[colnames(mcols(all_chr)) != "individual"]
  
  kpBars(gkp, data = genome_GR, y0 = rep(0, length(genome_GR)), y1 = rep(1, length(genome_GR)), col = "gray80", border = NA, r0 = ra, r1 = rz)
  kpAxis(gkp, side = 1, data.panel=1, cex = 0.7, r0 = ra, r1 = rz)
  kpText(gkp, chr=seqnames(genome_GR)[length(genome_GR)], x=width(genome_GR)[length(genome_GR)] + 2.6e7, y=0.5, labels = ind_lab, r0 = ra, r1 = rz)
  for (g in 1:length(groups)) {
    if (g == 1) {
      ga <- 0
      gz <- mcols(all_windows_large)[,groups[g]]
    } else {
      ga <- gz
      gz <- gz+mcols(all_windows_large)[,groups[g]]
    }
    kpBars(gkp, data = all_windows_large, y0 = ga, y1 = gz, col = group_col[groups[g]], border = NA, r0 = ra, r1 = rz)
  }
  # kpPlotRegions(gkp, data = all_windows_large, data.panel = "ideogram", col = "grey20", avoid.overlapping = T, r0 = ra, r1 = rz)
  # kpPlotRegions(gkp, data = all_windows_small, data.panel = "ideogram", col = "orange3", avoid.overlapping = T)
}
legend("right", #x = 0.9, y = 0.4, #"bottomright", #  yjust = .5, xjust = 0, # 
       title = "ancestry", legend = c(gsub("group_", "", names(group_col)), "undetermined"), 
       fill = c(group_col, "grey90"), bty = "n", xjust = 0, cex = 1.2)
dev.off()
```

```{r}
adm_blocks %>% 
  pivot_longer(cols = OB:ER, names_to = "ancestry", values_to = "proportion") %>% 
  mutate(start = start/1e6,
         end = end/1e6,
         individual = factor(individual, 
                             levels = adm_order)) %>% 
  group_by(chromosome, start, end, individual) %>% 
  mutate(proportion_min = cum_min(proportion),
         proportion_max = cum_max(proportion)) %>% 
  filter(individual %in% ind_3adm) %>% 
  ggplot() + 
  facet_grid(rows = vars(individual), cols = vars(chromosome),
             scale = "free_x", space = "free_x", switch = "x", shrink = F) + 
  geom_rect(aes(xmin = start, xmax = end,
                ymin = proportion_min, ymax = proportion_max,
                fill = ancestry),
            show.legend = F) +
  scale_fill_manual(values = group_col, na.value = "gray") +
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        axis.ticks = element_line(),
        axis.text.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        # strip.background.y = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.placement = "outside",
        strip.text.y = element_text(angle = 0, h = 0),
        panel.spacing.x = unit(4, "pt"), 
        panel.spacing.y = unit(0.5, "pt"), 
        panel.background = element_rect(fill = "gray60", color = "white")) +
  scale_y_continuous(breaks = 0.5) +
  scale_x_continuous(breaks = seq(0, 80, 10), expand = c(0, 0)) +
  xlab("Position (Mb)") + ylab("Ancestry assignment")
```

## averaged signals of admixed individuals



```{r}
adm_chr <- all_windows_large[!all_windows_large$individual %in% inds_ER]

avg_local <- do.call(rbind, lapply(seqlevels(adm_chr), function(ch) {
  print(ch)
  chr <- adm_chr[adm_chr@seqnames == ch]
  checkpoints <- sort(unique(c(start(chr), end(chr))))
  chr_score <- do.call(rbind, lapply(checkpoints, function(p) {
    print(paste(ch, ">", p))
    check_gr <- pintersect(chr, 
                           GRanges(ch, 
                                   IRanges(start = p, end = p), 
                                   strand = "+"))
    check_gr <- mcols(check_gr[check_gr$hit])
    check_score <- check_gr %>% 
      as.data.frame() %>% 
      select(- individual, - hit) %>% 
      summarise(across(is.numeric, ~ mean(.x, na.rm = TRUE))) %>% 
      mutate(chromosome = ch,
             position = p) %>% 
      relocate(chromosome, position)
    return(check_score)
  }))
  return(chr_score)
}))

avg_plot <- avg_local %>% 
  mutate(position = position/1e6,
         set = "") %>% 
  pivot_longer(cols = OB:ER, names_to = "ancestry", values_to = "proportion") %>% 
  ggplot() + 
  facet_grid(cols = vars(chromosome), scale = "free", space = "free", switch = "x") + 
  geom_line(aes(x = position, y = proportion, color = ancestry), show.legend = F) +
  scale_color_manual(values = group_col) +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(), 
        axis.ticks = element_line(),
        axis.title = element_text(size = 14),
        # panel.grid.major.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        strip.text.x = element_text(size = 14),
        strip.placement = "outside") +
  scale_x_continuous(breaks = seq(0, 80, 10)) +
  xlab("Position (Mb)") + ylab("Average proportion")
avg_plot
ggsave("./plots/averaged_admixture_signals.pdf", width = 15, height = 4)
```

## plot all individuals together

```{r}
cum_min <- function(x) {
  y <- x
  tt <- 0
  for (i in 1:length(x)) {
    if (x[i] == 0 || is.na(x[i])) {
      y[i] <- 0
    } else {
      y[i] <- tt
      tt  <- tt + x[i]
    }
  }
  return(y)
}

cum_min(c(0,0.5,0,0,0.5))

cum_max <- function(x) {
  y <- x
  tt <- 0
  for (i in 1:length(x)) {
    if (x[i] == 0 || is.na(x[i])) {
      y[i] <- 0
    } else {
      tt  <- tt + x[i]
      y[i] <- tt
    }
  }
  return(y)
}

cum_max(c(0,0.5,0.5,0,0))

```


```{r}
ind_order <- unique(c(inds_ER, inds_other[!inds_other %in% ind_3adm], 
                      inds_AG, inds_D, ind_3adm))


all_windows_large <- all_chr[width(all_chr) > 1e6,]
all_blocks <- mcols(all_windows_large) %>% 
  as.data.frame() %>% 
  mutate(chromosome = unlist(mapply(rep, 
                                    all_windows_large@seqnames@values, 
                                    all_windows_large@seqnames@lengths)),
         start = start(all_windows_large),
         end = end(all_windows_large)) %>% 
  relocate(chromosome, start, end, individual)

lai_plot <- all_blocks %>% 
  pivot_longer(cols = OB:ER, names_to = "ancestry", values_to = "proportion") %>% 
  mutate(start = start/1e6,
         end = end/1e6,
         individual = factor(individual, 
                             levels = ind_order)) %>% 
  group_by(chromosome, start, end, individual) %>% 
  mutate(proportion_min = cum_min(proportion),
         proportion_max = cum_max(proportion)) %>% 
  ggplot() + 
  facet_grid(rows = vars(individual), cols = vars(chromosome),
             scale = "free_x", space = "free_x", switch = "x", shrink = F) + 
  geom_rect(aes(xmin = start, xmax = end,
                ymin = proportion_min, ymax = proportion_max,
                fill = ancestry)) +
  scale_fill_manual(values = group_col, na.value = "gray") +
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        axis.ticks = element_line(),
        axis.text.y = element_blank(),
        axis.title = element_text(size = 14),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        # strip.background.y = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.placement = "outside",
        strip.text.y = element_text(angle = 0, h = 0),
        panel.spacing.x = unit(4, "pt"), 
        panel.spacing.y = unit(0.5, "pt"), 
        panel.background = element_rect(fill = "gray60", color = "white")) +
  scale_y_continuous(breaks = 0.5) +
  scale_x_continuous(breaks = seq(0, 80, 10), expand = c(0, 0))+
  xlab("position (Mb)") + ylab("ancestry assignment")
ggsave("./plots/all_lai.pdf", width = 20, height = 12)
```


```{r}
adm1 <- rownames(global)[global$AG > 0 & global$D == 0 & global$OB > 0]
adm2 <- rownames(global)[global$AG > 0 & global$D > 0 & global$OB == 0]
small_adm <- rownames(global)[!rownames(global) %in% c(inds_ER,inds_AG, inds_D, adm1, adm2)]
adm_order <- unique(c(small_adm, inds_AG, inds_D, adm1, adm2))

adm_blocks <- mcols(all_windows_large) %>% 
  as.data.frame() %>% 
  mutate(chromosome = unlist(mapply(rep, 
                                    all_windows_large@seqnames@values, 
                                    all_windows_large@seqnames@lengths)),
         start = start(all_windows_large),
         end = end(all_windows_large)) %>% 
  relocate(chromosome, start, end, individual) %>% 
  filter(individual %in% adm_order)

lai_amd_plot <- adm_blocks %>% 
  # filter(individual == "S-75") %>%
  pivot_longer(cols = OB:ER, names_to = "ancestry", values_to = "proportion") %>% 
  mutate(start = start/1e6,
         end = end/1e6,
         individual = factor(individual, 
                             levels = adm_order)) %>% 
  group_by(chromosome, start, end, individual) %>% 
  mutate(proportion_min = cum_min(proportion),
         proportion_max = cum_max(proportion)) %>% 
  ggplot() + 
  facet_grid(rows = vars(individual), cols = vars(chromosome),
             scale = "free_x", space = "free_x", switch = "x", shrink = F) + 
  geom_rect(aes(xmin = start, xmax = end,
                ymin = proportion_min, ymax = proportion_max,
                fill = ancestry)) +
  scale_fill_manual(values = group_col, na.value = "gray") +
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        axis.ticks = element_line(),
        axis.title.y = element_text(size = 14, vjust = 1),
        axis.text.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        # strip.background.y = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.placement = "outside",
        strip.text.y = element_text(size = 14, angle = 0, v = 0.5, h = 0),
        strip.text.x = element_text(size = 14),
        axis.title.x = element_text(size = 14, vjust = 1),
        panel.spacing.x = unit(4, "pt"), 
        panel.spacing.y = unit(0.5, "pt"), 
        panel.background = element_rect(fill = "gray60", color = "white")) +
  scale_y_continuous(breaks = 0.5) +
  scale_x_continuous(breaks = seq(0, 80, 10), expand = c(0, 0))+
  xlab("Position (Mb)") + ylab("Ancestry assignment")
```


```{r}
plot_list(gglist = list(lai_amd_plot, avg_plot), 
          ncol = 1, heights = c(3,1),
          labels = c("A", "B"))
```

## Admixture length


```{r}
all_blocks %>% 
  pivot_longer(cols = OB:ER, names_to = "ancestry", values_to = "proportion") %>% 
  mutate(individual = factor(individual, levels = ind_order),
         length = end - start + 1) %>% 
  filter(proportion != 0, !ancestry %in% c("C", "OB")) %>% 
  ggplot() +
  geom_density(aes(x = length/10^6, color = ancestry), linewidth = 1.5) +
  scale_color_manual(values = group_col[-3]) +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(), 
        axis.title = element_text(size = 16),
        # panel.grid.major.y = element_blank(),
        strip.background = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.placement = "outside") +
  scale_x_continuous(breaks = seq(0, 80, 10)) +
  xlab("Admixture block length (Mb)") + ylab("Density")
```



## recombination rate

```{r}
marey_chr_data <- "./snps_data/Coffea_canephora_Crouzillat2020.txt"
genmap <- read.table(marey_chr_data, header = T)
genmap$map <- sapply(genmap$map, function(i) 
  paste0("Chr ", sprintf("%02d", which(LETTERS == i))), 
  USE.NAMES = F)
```

```{r}
genmap_plot <- ggplot(genmap) + 
  geom_point(aes(x = phys/1e6, y = gen), size = 0.5) +
  facet_grid(cols = vars(map), scale = "free", space = "free", switch = "x") +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(), 
        axis.ticks = element_line(),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14),
        # panel.grid.major.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        strip.text.x = element_text(size = 14),
        strip.placement = "outside") +
  scale_x_continuous(breaks = seq(0, 80, 10)) +
  xlab("Position (Mb)") +
  ylab("Genetic distance (cM)")

plot_list(gglist = list(avg_plot, genmap_plot),
          ncol = 1, heights = c(0.5, 0.5),
          labels = c("A", "B"))
```

```{r}
genmap <- genmap %>% 
  mutate(rate = gen/phys*1e6) 
mean(genmap$rate)
nrow(genmap)
```

```{r}
marey_chr_data <- list.files("./snps_data", pattern = "chromosome", full.names = T)
rec_rate <- do.call(rbind, lapply(marey_chr_data, function(f) {
  df <- read.table(f, header = T)
  chr <- sprintf("%02d", which(gsub(".+chromosome|\\.txt", "", basename(f)) == LETTERS))
  df <- df %>% 
    mutate(chromosome = paste("Chr", chr))
  return(df)
  }))
mean(rec_rate$rec.rate, na.rm = T)
```

```{r}
ggplot(rec_rate) + 
  geom_line(aes(x = phys, y = rec.rate), size = 0.5) +
  facet_grid(cols = vars(chromosome), scale = "free", space = "free", switch = "x") +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(), 
        axis.ticks = element_line(),
        axis.title.x = element_text(size = 12),
        axis.title.y.left = element_text(size = 12),
        axis.title.y.right = element_text(size = 12, color = "gray50"),
        # panel.grid.major.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        strip.text.y = element_text(angle = 0, h = 0),
        strip.placement = "outside") +
  scale_x_continuous(breaks = seq(0, 80, 10)) +
  xlab("Position (Mb)") +
  ylab("Recombination rate (cM/Mb)")
```

recom rate + lai
```{r}
avg_local_long <- avg_local %>% 
  mutate(position = position/1e6,
         set = "") %>% 
  pivot_longer(cols = OB:ER, names_to = "ancestry", values_to = "proportion") 

genmap_lai_plot <- ggplot(avg_local_long) + 
  facet_grid(cols = vars(chromosome), scale = "free", space = "free", switch = "x") +
  # facet_wrap(vars(chromosome), nrow = 1, scale = "free_x", strip.position = "bottom") +
  geom_point(data = recom %>% rename(chromosome = map) %>% 
               mutate(set = ""), 
             aes(x = phys/1e6, y = gen/250), 
             size = 0.8, color = "gray65") +
  geom_line(aes(x = position, y = proportion, color = ancestry), 
            size = 0.8, show.legend = F) +
  scale_y_continuous(name = "Average ancestry proportion",
                     sec.axis = sec_axis(trans = ~.*250, name = "Genetic distance (cM)")) +
  scale_color_manual(values = group_col) +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(), 
        axis.ticks = element_line(),
        axis.title.x = element_text(size = 12),
        axis.title.y.left = element_text(size = 12),
        axis.title.y.right = element_text(size = 12, color = "gray50"),
        # panel.grid.major.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        strip.text.y = element_text(angle = 0, h = 0),
        strip.placement = "outside") +
  scale_x_continuous(breaks = seq(0, 80, 10)) +
  xlab("Position (Mb)") 

tiff(file.path(plot_dir, "lai_recom.tiff"),
     width = 15, heigh = 10, unit = "in", res = 1000)
plot_list(gglist = list(lai_amd_plot, genmap_lai_plot), 
          ncol = 1, heights = c(4,1.5),
          labels = c("A", "B"))
dev.off()
```

## adaptive introgression (Fst)
### Fst
Estimate Fst by windows of 5kb between VN hybrids and the donor group AG (7 inds), D (9 inds), ER (12 inds) (threshold 0.7)

```{r}
## all hybrids together
vn_hybrids <- data.frame(Label = rownames(global[!rownames(global) %in% inds_ER,]),
                         assign = "hybrid") %>% 
  mutate(Label = paste0(Label, ifelse(grepl("TR\\d", Label), "m", "")))
write.table(vn_hybrids, "./snp_stats/vn_hybrids.txt", sep = "\t", 
            row.names = F, col.names = F, quote = F)

## only hybrids with AG admixture (include 4 three-way admixed inds)
vn_AG <- data.frame(Label = rownames(global[rownames(global) %in% c(inds_AG),]),
                         assign = "hybrid_AG") %>% 
  mutate(Label = paste0(Label, ifelse(grepl("TR\\d", Label), "m", "")))
write.table(vn_AG, "./snp_stats/vn_hybrids_AG.txt", sep = "\t", 
            row.names = F, col.names = F, quote = F)

## only hybrids with D admixture (include 4 three-way admixed inds except for S-122)
vn_D <- data.frame(Label = rownames(global[rownames(global) %in% c(inds_D),]),
                         assign = "hybrid_D") %>% 
  mutate(Label = paste0(Label, ifelse(grepl("TR\\d", Label), "m", "")))
write.table(vn_D, "./snp_stats/vn_hybrids_D.txt", sep = "\t", 
            row.names = F, col.names = F, quote = F)

## ancestral groups (AG, D, ER)
ref_inds <- read.csv("./snp_stats/af_samples_ancestry.csv")
af_donor <- ref_inds %>% 
  dplyr::filter(assign %in% c("D", "AG", "ER")) %>% 
  dplyr::select(Label, assign) 
for (i in unique(af_donor$assign)) {
  write.table(fst_pop %>% dplyr::filter(assign == i), 
              paste0("./snp_stats/donor_", i, ".txt"), 
              sep = "\t", 
              row.names = F, col.names = F, quote = F)
}

write.table(fst_pop, "./snp_stats/snmf_thin_pop.txt", sep = "\t", 
            row.names = F, col.names = F, quote = F)
```

```{bash}
sbatch /shared/ifbstor1/projects/elai_most/diversity_wasi/seq_analysis/snp_stats/snp_stats.sh
```

```{r}
all_windows_large <- all_chr[width(all_chr) > 1e6,]
all_blocks <- mcols(all_windows_large) %>% 
  as.data.frame() %>% 
  mutate(chromosome = unlist(mapply(rep, 
                                    all_windows_large@seqnames@values, 
                                    all_windows_large@seqnames@lengths)),
         start = start(all_windows_large),
         end = end(all_windows_large)) %>% 
  relocate(chromosome, start, end, individual)
```

```{r}
hybrid_AG <- read.delim("./snp_stats/hybrid_AG.windowed.weir.fst")

hybrid_AG %>% 
  mutate(position = BIN_START/1e6,
         set = "") %>% 
  ggplot(.) + 
  facet_grid(cols = vars(CHROM), rows = vars(set), scale = "free", space = "free", switch = "x") + 
  geom_line(aes(x = position, y = MEAN_FST), color = group_col["AG"], show.legend = F) +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(), 
        axis.ticks = element_line(),
        # panel.grid.major.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        strip.text.y = element_text(angle = 0, h = 0),
        strip.placement = "outside") +
  scale_x_continuous(breaks = seq(0, 80, 10)) +
  scale_y_continuous(limits = c(0, 0.5), breaks = seq(0, .50, .10)) +
  xlab("Position (Mb)") + ylab("Fst between hybrids and group AG")
```

```{r}
hybrid_D <- read.delim("./snp_stats/hybrid_D.windowed.weir.fst")

hybrid_D %>% 
  mutate(position = BIN_START/1e6,
         set = "") %>% 
  ggplot(.) + 
  facet_grid(cols = vars(CHROM), rows = vars(set), scale = "free", space = "free", switch = "x") + 
  geom_line(aes(x = position, y = MEAN_FST), color = group_col["D"], show.legend = F) +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(), 
        axis.ticks = element_line(),
        # panel.grid.major.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        strip.text.y = element_text(angle = 0, h = 0),
        strip.placement = "outside") +
  scale_x_continuous(breaks = seq(0, 80, 10)) +
  scale_y_continuous(limits = c(0, 0.5), breaks = seq(0, .50, .10)) +
  xlab("Position (Mb)") + ylab("Fst between hybrids and group D")
```

```{r}
hybrid_ER <- read.delim("./snp_stats/hybrid_D.windowed.weir.fst")

do.call(rbind, list(hybrid_ER %>% mutate(group = "ER"), 
                 hybrid_AG %>% mutate(group = "AG"), 
                 hybrid_D %>% mutate(group = "D"))) %>% 
  as.data.frame() %>% 
# hybrid_D %>% 
  mutate(position = BIN_START/1e6) %>% 
  ggplot(.) + 
  facet_grid(cols = vars(CHROM), rows = vars(group), scale = "free", space = "free", switch = "both") + 
  geom_line(aes(x = position, y = MEAN_FST, color = group), show.legend = F) +
  scale_color_manual(values = group_col[c("AG", "D", "ER")]) +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(), 
        axis.ticks = element_line(),
        # panel.grid.major.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        strip.text.y = element_text(angle = 0, h = 0.5, size = 12),
        strip.placement = "outside") +
  scale_x_continuous(breaks = seq(0, 80, 10)) +
  scale_y_continuous(limits = c(0, 0.5), breaks = seq(0, .50, .10)) +
  xlab("Position (Mb)") + ylab("Fst between hybrids and ancestral groups")
```



```{r}
hybridAG_donorAG <- read.delim("./snp_stats/hybridAG_donorAG.windowed.weir.fst")
hybridAG_donorER <- read.delim("./snp_stats/hybridAG_donorER.windowed.weir.fst")
donorAG_donorER <- read.delim("./snp_stats/donorAG_donorER.windowed.weir.fst")

hybridAG_fst <- rbind(hybridAG_donorER %>% mutate(group = "ER"),
                      hybridAG_donorAG %>% mutate(group = "AG"))

hybridAG_plot <- all_blocks %>% 
  filter(individual %in% c(inds_AG)) %>% 
  pivot_longer(cols = OB:ER, names_to = "ancestry", values_to = "proportion") %>% 
  mutate(start = start/1e6,
         end = end/1e6,
         individual = factor(individual, 
                             levels = ind_order)) %>% 
  group_by(chromosome, start, end, individual) %>% 
  mutate(proportion_min = cum_min(proportion),
         proportion_max = cum_max(proportion)) %>% 
  ggplot() + 
  facet_grid(rows = vars(individual), cols = vars(chromosome),
             scale = "free_x", space = "free_x", switch = "x", shrink = F) + 
  geom_rect(aes(xmin = start, xmax = end,
                ymin = proportion_min, ymax = proportion_max,
                fill = ancestry)) +
  scale_fill_manual(values = group_col, na.value = "gray") +
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        axis.ticks = element_line(),
        axis.text.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        # strip.background.y = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.placement = "outside",
        strip.text.y = element_text(angle = 0, h = 0, size = 11),
        panel.spacing.x = unit(4, "pt"), 
        panel.spacing.y = unit(0.5, "pt"), 
        panel.background = element_rect(fill = "gray60", color = "white")) +
  scale_y_continuous(breaks = 0.5) +
  scale_x_continuous(breaks = seq(0, 80, 10), expand = c(0, 0))+
  xlab("Position (Mb)") + ylab("Ancestry assignment")

hybridAG_fst_plot <- hybridAG_fst %>%
  mutate(group = factor(group, levels = c("ER", "AG"))) %>% 
  mutate(position = BIN_START/1e6) %>% 
  ggplot(.) + 
  facet_grid(cols = vars(CHROM), scale = "free", space = "free", switch = "x") + 
  geom_line(aes(x = position, y = MEAN_FST, color = group)) +
  scale_color_manual(values = group_col[c("AG", "ER")]) +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(), 
        axis.ticks = element_line(),
        # panel.grid.major.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        strip.text.y = element_text(angle = 0, h = 0.5, size = 12),
        strip.placement = "outside") +
  scale_x_continuous(breaks = seq(0, 80, 10)) +
  scale_y_continuous(limits = c(0, 0.5), breaks = seq(0, .50, .10)) +
  xlab("Position (Mb)") + ylab("Fst. AG admixed vs ancestral")

donorAG_donorER %>%
  mutate(position = BIN_START/1e6) %>% 
  ggplot(.) + 
  facet_grid(cols = vars(CHROM), scale = "free", space = "free", switch = "x") + 
  geom_line(aes(x = position, y = MEAN_FST), color = "gray") +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(), 
        axis.ticks = element_line(),
        # panel.grid.major.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        strip.text.y = element_text(angle = 0, h = 0.5, size = 12),
        strip.placement = "outside") +
  scale_x_continuous(breaks = seq(0, 80, 10)) +
  scale_y_continuous(limits = c(0, 0.5), breaks = seq(0, .50, .10)) +
  xlab("Position (Mb)") + ylab("Fst. AG vs ER")

plot_list(gglist = list(hybridAG_plot, hybridAG_fst_plot), 
          ncol = 1, heights = c(.6,1),
          labels = c("A", "B"))
```

```{r}
hybridD_donorD <- read.delim("./snp_stats/hybridD_donorD.windowed.weir.fst")
hybridD_donorER <- read.delim("./snp_stats/hybridD_donorER.windowed.weir.fst")
donorD_donorER <- read.delim("./snp_stats/donorD_donorER.windowed.weir.fst")

hybridD_fst <- rbind(hybridD_donorD %>% mutate(group = "D"),
                      hybridD_donorER %>% mutate(group = "ER"))

hybridD_plot <- all_blocks %>% 
  filter(individual %in% c(inds_D)) %>% 
  pivot_longer(cols = OB:ER, names_to = "ancestry", values_to = "proportion") %>% 
  mutate(start = start/1e6,
         end = end/1e6,
         individual = factor(individual, 
                             levels = ind_order)) %>% 
  group_by(chromosome, start, end, individual) %>% 
  mutate(proportion_min = cum_min(proportion),
         proportion_max = cum_max(proportion)) %>% 
  ggplot() + 
  facet_grid(rows = vars(individual), cols = vars(chromosome),
             scale = "free_x", space = "free_x", switch = "x", shrink = F) + 
  geom_rect(aes(xmin = start, xmax = end,
                ymin = proportion_min, ymax = proportion_max,
                fill = ancestry)) +
  scale_fill_manual(values = group_col, na.value = "gray") +
  theme_minimal() + 
  theme(panel.grid = element_blank(), 
        axis.ticks = element_line(),
        axis.text.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        # strip.background.y = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.placement = "outside",
        strip.text.y = element_text(angle = 0, h = 0, size = 11),
        panel.spacing.x = unit(4, "pt"), 
        panel.spacing.y = unit(0.5, "pt"), 
        panel.background = element_rect(fill = "gray60", color = "white")) +
  scale_y_continuous(breaks = 0.5) +
  scale_x_continuous(breaks = seq(0, 80, 10), expand = c(0, 0))+
  xlab("Position (Mb)") + ylab("Ancestry assignment")

hybridD_fst_plot <- hybridD_fst %>%
  mutate(position = BIN_START/1e6) %>% 
  ggplot(.) + 
  facet_grid(cols = vars(CHROM), scale = "free", space = "free", switch = "x") + 
  geom_line(aes(x = position, y = MEAN_FST, color = group)) +
  scale_color_manual(values = group_col[c("D", "ER")]) +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(), 
        axis.ticks = element_line(),
        # panel.grid.major.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        strip.text.y = element_text(angle = 0, h = 0.5, size = 12),
        strip.placement = "outside") +
  scale_x_continuous(breaks = seq(0, 80, 10)) +
  scale_y_continuous(limits = c(0, 0.5), breaks = seq(0, .50, .10)) +
  xlab("Position (Mb)") + ylab("Fst. D admixed vs ancestral")

donorD_donorER %>%
  mutate(position = BIN_START/1e6) %>% 
  ggplot(.) + 
  facet_grid(cols = vars(CHROM), scale = "free", space = "free", switch = "x") + 
  geom_line(aes(x = position, y = MEAN_FST), color = "gray") +
  scale_color_manual(values = group_col[c("D", "ER")]) +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(), 
        axis.ticks = element_line(),
        # panel.grid.major.y = element_blank(),
        strip.background.x = element_rect(fill = "#dadde6", color = "#dadde6"),
        strip.background.y = element_rect(fill = NA, color = NA),
        strip.text.y = element_text(angle = 0, h = 0.5, size = 12),
        strip.placement = "outside") +
  scale_x_continuous(breaks = seq(0, 80, 10)) +
  scale_y_continuous(limits = c(0, 0.5), breaks = seq(0, .50, .10)) +
  xlab("Position (Mb)") + ylab("Fst. D vs ER")

plot_list(gglist = list(hybridD_plot, hybridD_fst_plot), 
          ncol = 1, heights = c(1,1),
          labels = c("A", "B"))
```


### distribution of ancestry proportion

```{r}
avg_local %>% 
  pivot_longer(., cols = OB:ER, names_to = "group", values_to = "proportion") %>% 
  filter(!group %in% c("C", "OB")) %>% 
  ggplot() +
  facet_grid(rows = vars(group), scales = "fixed", space = "fixed") +
  geom_histogram(aes(x = proportion, fill = group), binwidth = .01, show.legend = F) +
  # geom_density(aes(x = proportion, color = group)) +
  scale_color_manual(values = group_col) +
  scale_fill_manual(values = group_col) +
  theme_minimal() + 
  theme(panel.border = element_rect(color = "gray", fill = NA),
        strip.text = element_text(size = 16),
        axis.title = element_text(size = 16)) +
  scale_y_continuous(expand = c(0,0)) +
  ylab("Number of SNPs") +
  xlab("Average ancestry proportion at a SNP")
```

```{r}
window_lai <- avg_local %>% 
  pivot_longer(., cols = OB:ER, names_to = "group", values_to = "proportion") %>% 
  group_by(chromosome, group) %>% 
  mutate(window = cut(position, breaks = seq(0, 70*10^6, 50000))) %>% 
  mutate(window = as.numeric(gsub("\\(.+\\,|]", "", as.character(window)))/10^6) %>% 
  group_by(chromosome, group, window) %>% 
  summarise(mean_prop = mean(proportion))


left_join(window_lai, rec_rate %>% rename(window = phys), 
          by = c("chromosome", "window")) %>% 
  filter(!group %in%  c("C", "OB")) %>% 
  ggplot() +
  facet_grid(rows = vars(group)) +
  geom_point(aes(x = rec.rate, y = mean_prop, color = group), show.legend = F) +
  scale_color_manual(values = group_col) +
  scale_fill_manual(values = group_col) +
  theme_minimal() + 
  theme(panel.border = element_rect(color = "gray", fill = NA),
        strip.text = element_text(size = 16)) +
  xlab("Mean recombination rate in 50 kb windows") + 
  ylab("Mean ancestry proportion in 50 kb windows")

```


## for asic poster
lai plots of 3 individuals

```{r}
pp3 <- getDefaultPlotParams(plot.type=6)
pp3$topmargin <- 5
pp3$bottommargin <- 5
pp3$data1outmargin <- 20

chr_ind <- all_chr[all_chr$individual == "TR15"]
all_windows_large <- chr_ind[width(chr_ind) > 1e6,]
all_windows_small <- chr_ind[width(chr_ind) < 1e6,]
groups <- colnames(mcols(all_chr))[colnames(mcols(all_chr)) != "individual"]
gkp <- plotKaryotype(genome = genome_GR, plot.type = 6,
                     plot.params = pp3) 
kpDataBackground(gkp, data.panel ="ideogram", color = "grey70")
# kpAddBaseNumbers(gkp, tick.dist = 1e7)
for (g in 1:length(groups)) {
  if (g == 1) {
    ga <- 0
    gz <- mcols(all_windows_large)[,groups[g]]
  } else {
    ga <- gz
    gz <- gz+mcols(all_windows_large)[,groups[g]]
  }
  kpBars(gkp, data = all_windows_large, y0 = ga, y1 = gz, col = group_col[groups[g]], border = NA)
}
# kpAxis(gkp, side = 2, data.panel=1, cex = 0.7)
```


```{r}
chr_ind <- all_chr[all_chr$individual == "S-75"]
all_windows_large <- chr_ind[width(chr_ind) > 1e6,]
all_windows_small <- chr_ind[width(chr_ind) < 1e6,]
groups <- colnames(mcols(all_chr))[colnames(mcols(all_chr)) != "individual"]
gkp <- plotKaryotype(genome = genome_GR, plot.type = 6,
                     plot.params = pp3) 
kpDataBackground(gkp, data.panel ="ideogram", color = "grey70")
# kpAddBaseNumbers(gkp, tick.dist = 1e7)
for (g in 1:length(groups)) {
  if (g == 1) {
    ga <- 0
    gz <- mcols(all_windows_large)[,groups[g]]
  } else {
    ga <- gz
    gz <- gz+mcols(all_windows_large)[,groups[g]]
  }
  kpBars(gkp, data = all_windows_large, y0 = ga, y1 = gz, col = group_col[groups[g]], border = NA)
}
# kpAxis(gkp, side = 2, data.panel=1, cex = 0.7)
```

```{r}
chr_ind <- all_chr[all_chr$individual == "S-122"]
all_windows_large <- chr_ind[width(chr_ind) > 1e6,]
all_windows_small <- chr_ind[width(chr_ind) < 1e6,]
groups <- colnames(mcols(all_chr))[colnames(mcols(all_chr)) != "individual"]
gkp <- plotKaryotype(genome = genome_GR, plot.type = 6,
                     plot.params = pp3) 
kpDataBackground(gkp, data.panel ="ideogram", color = "grey70")
# kpAddBaseNumbers(gkp, tick.dist = 1e7)
for (g in 1:length(groups)) {
  if (g == 1) {
    ga <- 0
    gz <- mcols(all_windows_large)[,groups[g]]
  } else {
    ga <- gz
    gz <- gz+mcols(all_windows_large)[,groups[g]]
  }
  kpBars(gkp, data = all_windows_large, y0 = ga, y1 = gz, col = group_col[groups[g]], border = NA)
}
# kpAxis(gkp, side = 2, data.panel=1, cex = 0.7)
```

